Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model

### 2023-03-19 src/docs_paper/paper.Rmd
Why MCMC is slow (could be better explained)

The practical requirements for the model, combined with its relative complexity, present a difficult Bayesian inference problem.
Any inferential strategy must be fast enough for interactive review and iteration of modelling results, as well as easy to run in production by country teams, ruling out prohibitively slow Markov chain Monte Carlo (MCMC) approaches.
Inference is currently conducted using an empirical Bayes (EB) approach, with a Gaussian approximation to the latent field, via the Template Model Builder (`TMB`) \textsf{R} package [@kristensen2016tmb].
Owing to its speed and flexibility, `TMB` has recently been gaining popularity more broadly in spatial statistics [@osgood2022statistical].
Inference in `TMB` is based on optimisation of a \textsf{C++} template function, with the option available to use a Laplace approximation to integrate out any subset of the function arguments, which for the Naomi model, we use for the high-dimensional latent field parameters.
Taking inspiration from the AD Model Builder (ADMB) package [@fournier2012ad], `TMB` uses automatic differentiation [@baydin2017automatic] to calculate the derivatives required for numerical optimisation routines and the Laplace approximation.
Although this approach has favourable computational properties, it would be preferable to account more fully for hyperparameter uncertainty than is possible in an empirical Bayes framework.
This has motivated us to look for an approach closer to full Bayesian inference, which is also flexible enough to be compatible with the model, as well as fast enough to be run in production by country teams.

@@

Other work aiming to extend the scope of the INLA method includes the `inlabru` \textsc{R} package of @bachl2019inlabru, and the INLA within MCMC approach of @gomez2018markov, both of which leverage the `R-INLA` \textsc{R} package [@martins2013bayesian].
To approximate non-linear predictors, `inlabru` makes iterated use of `R-INLA` using linearisation, extending the scope of INLA to LGMs with some suitably small amount of non-linearity.
<!-- See https://inlabru-org.github.io/inlabru/articles/method.html -->
<!-- Want to say in particular what the class of models that `inlabru` allows you to fit are -->
INLA within MCMC is suitable for models which are LGMs, though only conditional on some subset of the parameters being fixed.
For these models, Metropolis-Hastings style algorithms be defined to update the fixed parameters, with acceptance probabilities calculated using by repeat calls to `R-INLA`.
It would or would not be possible to fit the Naomi model in these frameworks.

>>>>

The practical requirements for the model, combined with its relative complexity, present a difficult Bayesian inference problem.
Any inferential strategy must be fast enough for interactive review and iteration of modelling results, as well as easy to run in production by country teams.
Markov chain Monte Carlo (MCMC) approaches are prohibitively slow due to the scale of the model and challenging features of its posterior geometry [@neal2003slice].
Inference is currently conducted using an empirical Bayes (EB) approach, with a Gaussian approximation to the latent field, via the Template Model Builder (`TMB`) \textsf{R} package [@kristensen2016tmb].
Owing to its speed and flexibility, `TMB` has recently been gaining popularity more broadly in spatial statistics [@osgood2022statistical].
Inference in `TMB` is based on optimisation of a \textsf{C++} template function, with the option available to use a Laplace approximation to integrate out any subset of the function arguments, which for the Naomi model, we use for the high-dimensional latent field parameters.
Taking inspiration from the AD Model Builder (ADMB) package [@fournier2012ad], `TMB` uses automatic differentiation [@baydin2017automatic] to calculate the derivatives required for numerical optimisation routines and the Laplace approximation.
Although this approach has favourable computational properties, it would be preferable to account more fully for hyperparameter uncertainty than is possible in an empirical Bayes framework.
This has motivated us to look for an approach closer to full Bayesian inference, which is also flexible enough to be compatible with the model, as well as fast enough to be run in production by country teams.

@@

Other work aiming to extend the scope of the INLA method includes the `inlabru` \textsc{R} package of @bachl2019inlabru, and the INLA within MCMC approach of @gomez2018markov, both of which leverage the `R-INLA` \textsc{R} package [@martins2013bayesian].
To approximate non-linear predictors, `inlabru` makes iterated use of `R-INLA` using linearisation, extending the scope of INLA to LGMs with some suitably small amount of non-linearity.
<!-- See https://inlabru-org.github.io/inlabru/articles/method.html -->
<!-- Want to say in particular what the class of models that `inlabru` allows you to fit are -->
INLA within MCMC is suitable for models which are LGMs, though only conditional on some subset of the parameters being fixed.
For these models, Metropolis-Hastings style algorithms be defined to update the fixed parameters, with acceptance probabilities calculated using by repeat calls to `R-INLA`.
<!-- It would or would not be possible to fit the Naomi model in these frameworks. -->

<!-- Note: can fit hyperparameters with Laplace on latent field using HMC via tmbstan::tmbstan(laplace = TRUE) -->
